### 2023-01-23 template.tex
sorted out the cls issue

% JuliaCon proceedings template
\documentclass{juliacon}
\setcounter{page}{1}
\begin{document}
\hypertarget{sec-intro}{%
\section{Introduction}\label{sec-intro}}

\emph{TODO} Create a template that demonstrates the appearance,
formatting, layout, and functionality of your format. Learn more about
journal formats at \url{https://quarto.org/docs/journals/}.

\hypertarget{proof-of-concept}{%
\subsection{Proof-of-Concept}\label{proof-of-concept}}

This is a proof-of-concept for how we could use Quarto for JuliaCon
proceedings. For current submissions, please ignore this repo and follow
the official instructions
\href{https://github.com/JuliaCon/JuliaConSubmission.jl}{here}
\end{document}