### 2023-11-14 template.qmd
basically done

---
title: Juliacon Proceedings in Quarto
format:
  juliacon-proceedings-pdf:
    keep-tex: true  
  juliacon-proceedings-html: default
author:
  - name: Patrick Altmeyer
    affiliations:
      - Delft University of Technology
    orcid: 0000-0003-4726-8613
    email: p.altmeyer@tudelft.nl
    url: https://www.paltmeyer.com/
  - name: Cynthia C. S. Liem
    affiliations:
      - Delft University of Technology
abstract: |
  This document is a Quarto template demonstrating the `juliacon-proceedings` format.
year: 2022
keywords: [Template, Demo, Quarto, JuliaCon]
bibliography: bibliography.bib  
---

# Introduction {#sec-intro}

*TODO* Create a template that demonstrates the appearance, formatting, layout, and functionality of your format. Learn more about journal formats at <https://quarto.org/docs/journals/>.

## Proof-of-Concept

This is a proof-of-concept for how we could use Quarto for JuliaCon proceedings. For current submissions, please ignore this repo and follow the official instructions [here](https://github.com/JuliaCon/JuliaConSubmission.jl).

## Code

```{.julia}
println("Hello, world!")
```