### 2024-01-08 template.qmd
going to turn the presentation into a website

---
title: JuliaCon Proceedings in Quarto
format:
  juliacon-proceedings-pdf:
    keep-tex: true  
  juliacon-proceedings-html: default
author:
  - name: Patrick Altmeyer
    affiliations:
      - Delft University of Technology
    orcid: 0000-0003-4726-8613
    email: p.altmeyer@tudelft.nl
    url: https://www.paltmeyer.com/
abstract: |
  This document is a template demonstrating the `juliacon-proceedings` format. It is a standard Quarto document composed of Markdown, LaTeX and code chunks. It is rendered to PDF and HTML. The PDF is rendered using the `juliacon-proceedings-pdf` format, which is based on the `juliacon-proceedings` LaTeX class. The HTML is rendered using the `juliacon-proceedings-html` format, which is based on the `juliacon-proceedings` HTML template.
keywords: [Template, Demo, Quarto, JuliaCon]
bibliography: bibliography.bib  
jupyter: julia-1.9
---

```{julia}
#| echo: false
#| output: false

using Pkg; Pkg.activate(".")
```

# Introduction {#sec-intro}

This is a template for writing a JuliaCon Proceedings article in Quarto. This is a proof-of-concept for how we could use Quarto for JuliaCon proceedings. For current submissions, please ignore this repo and follow the official instructions [here](https://github.com/JuliaCon/JuliaConSubmission.jl).

## What is Quarto?

[Quarto](https://quarto.org/) makes it easy to write reproducible documents that can be rendered to PDF, HTML, Word and more. It is based on Markdown, which is easy to learn and write. It also supports LaTeX, which is useful for more advanced formatting. As this extension demonstrates, Quarto is also very flexible and can be extended with custom templates and styles.

## Why Quarto?

By embracing Quarto, JuliaCon Proceedings can set an example for how to write reproducible documents. We would not only make it easier for authors to write their submissions but also open the door for more advanced features such as interactive figures and executable code blocks in HTML documents.

# About this template

This template is based on the existing [JuliaCon Proceedings LaTeX template](https://github.com/JuliaCon/JuliaConSubmission.jl). The rendered versions can be found here: [pdf](https://www.paltmeyer.com/quarto-juliacon-proceedings/template.pdf), [html](https://www.paltmeyer.com/quarto-juliacon-proceedings/template.html).

## Basic Usage

To use the Quarto [extension](https://github.com/pat-alt/quarto-juliacon-proceedings) that provides this template, you can create a new project as follows:

```
quarto use template pat-alt/quarto-juliacon-proceedings
```

Alternatively, you can add the extension to an existing project:

```
quarto add pat-alt/quarto-juliacon-proceedings
```

Then, add the format to your document options:

```yaml
format:
  juliacon-proceedings-pdf: default
```

# Code

There are various options to present code using this template. We could leverage Quarto's existing support for executable, cross-referenceable code. Alternatively, we could use the existing `lstlisting` environment for Julia code.

## Executable Listings  (version $\geq$ 1.4.0) {#sec-executable-listings}

Relying on Quarto's support for executable listings is the most flexible option as it allows us to use all of Quarto's features. It comes with numerous advantages:

1. We can use the same code chunks for both PDF and HTML output (and more) with consistent formatting.
2. Using executable code ensures that things stay up-to-date and potential errors in the code are identified early: as you render your document, the code will be executed and the output will be inserted into the document.
3. The output of executable code can be cross-referenced. For example, we can reference a figure that is generated by a code chunk and the reference will be updated automatically when the figure number changes.
4. Code chunks can also be hidden. In HTML, there is additional support for code folding.
5. The code itself can be cross-referenced.

From the [Quarto Docs](https://quarto.org/docs/prerelease/1.4/crossref.html#cross-referenceable-listings-of-executable-code-blocks):

> To create cross-referenceable code listings from executable code blocks, use `lst-label` and `lst-cap`. 
> --- [Quarto Docs](https://quarto.org/docs/prerelease/1.4/crossref.html#cross-referenceable-listings-of-executable-code-blocks)

For example, the following code block will be labelled as `lst-1` and calling `@lst-1` will render as @lst-1. The figure it generates will be labelled as `fig-1` and calling `@fig-1` will render as @fig-1.

```{julia}
#| lst-label: lst-1
#| lst-cap: A listing caption.
#| label: fig-1
#| fig-cap: A figure caption.
#| eval: true
#| echo: true

using Plots

x = -3.0:0.01:3.0
y = rand(length(x))
plot(x, y)
```

::: {.callout-note}

# Current Limitations

- Support for executable listings is only available in the latest version of Quarto (version $\geq$ 1.4.0).
- It is currently not clear how to support custom formatting, for example through the existing `jlcode.sty` environment. 
- Cross-referencing listings does not seem to work in HTML output.

:::

Alternatively, we can use the existing `lstlisting` environment for Julia code using the `julia` language option. This is the same environment that is used for Julia code in the existing LaTeX template. Under the hood, this uses a lua filter to convert the code to LaTeX. The code creates @fig-2 which can be cross-referenced as `@fig-2`.

```{julia}
#| label: fig-2
#| fig-cap: A figure caption.
#| eval: true
#| echo: true

using Plots

x = -3.0:0.01:3.0
y = cumsum(rand(length(x)))
plot(x, y)
```

::: {.callout-note}

# Current Limitations

I don't know how to cross-reference the code itself.

:::

## Static Listings

The special environment that is already defined for Julia code is still compatible with this template.

\begin{verbatim}
\begin{lstlisting}[
    language = Julia, 
    numbers=left, 
    label={lst:exmplg}, 
    caption={Example Code Block.}
]
using Plots

x = -3.0:0.01:3.0
y = rand(length(x))
plot(x, y)
\end{lstlisting}
\end{verbatim}
\begin{lstlisting}[
    language = Julia, 
    numbers=left, 
    label={lst:exmplg}, 
    caption={Example Code Block.}
]
using Plots

x = -3.0:0.01:3.0
y = rand(length(x))
plot(x, y)
\end{lstlisting}

# Tables

As with code, there are various options to present tables. Using standard Markdown syntax is the most flexible option because it is compatible with various output formats. Alternatively, standard LaTeX syntax can be used.

## Markdown Tables

Markdown tables are rendered to LaTeX using the `booktabs` package. This is the recommended way to create tables because it is compatible with various output formats. Tables can be cross-referenced as `@tbl-1` which renders as @tbl-1.

```

| Default | Left | Right | Center |
|---------|:-----|------:|:------:|
| 12      | 12   |    12 |   12   |
| 123     | 123  |   123 |  123   |
| 1       | 1    |     1 |   1    |

: Demonstration of pipe table syntax {#tbl-1}

```

| Default | Left | Right | Center |
|---------|:-----|------:|:------:|
| 12      | 12   |    12 |   12   |
| 123     | 123  |   123 |  123   |
| 1       | 1    |     1 |   1    |

: Demonstration of pipe table syntax {#tbl-1}

::: {.callout-note}

# Current Limitations

This seems to currently mess with executable listing (@sec-executable-listings).

:::

## LaTeX Tables

Standard LaTeX syntax can also be used to create tables. To use standard Quarto cross-referencing, the LaTeX syntax needs to be wrapped in a `div` with the `tbl` class. The table caption can be added as a paragraph. The following creates @tbl-2 which can be cross-referenced as `@tbl-2`.

```
::: {#tbl-2}

\begin{tabular}{|l|l|}\hline
Age & Frequency \\ \hline
18--25  & 15 \\
26--35  & 33 \\
36--45  & 22 \\ \hline
\end{tabular}

This is a table caption.

:::
```

::: {#tbl-2}

\begin{tabular}{|l|l|}\hline
Age & Frequency \\ \hline
18--25  & 15 \\
26--35  & 33 \\
36--45  & 22 \\ \hline
\end{tabular}

This is a table caption.

:::

# Citations

This `@altmeyer2023endogenous` renders the citation as @altmeyer2023endogenous. The bibliography is defined in `bibliography.bib`.

# References